



```{r initialize, echo=FALSE,results='hide',message=FALSE,warning=FALSE}
# Initialize!

# Stop R from interpreting strings as factors. 
# When you want factors, you'll know it.
options(stringsAsFactors = FALSE)

# Load up the requisite packages
library(knitr)     # It's a knitr document, after all...
library(to1check)  # All TO1-related functions
library(ggplot2)
library(scales)
library(gridExtra)
library(RColorBrewer)
library(plyr)      # For the count() and arrange() functions
library(mpmisc)    # For printing data.frames nicely
library(lubridate) # For calculating number of weeks remaining

# Set default chunk options
opts_chunk$set(echo = FALSE,
               results = 'asis',
               message = FALSE,
               warning = FALSE,
               error = TRUE,
               fig.width = 10,
               out.width = '800px',
               dpi = 600)



# Set up some colors to use throughout
colorwheel <- c( "#F03B20", "#31A354", "#1F78B4")

# Load the ggplot2 theme
source(file.path("..", "css", "theme_tbesc.r"))


# Load the local info (edit this if your results are wrong)
local_facts <- read.csv("local_facts.csv")

# Load the latest cleaned data
load(local_facts$datapath)

# Set aside those who were successfully enrolled - we'll generally disregard 
# those who declined or who didn't complete enrollment 

enrolled <- with(to1clean$master,
    StudyID[!CloseReason %in% c("Didn't complete enrollment", 
                                "Withdrew", 
                                "Not eligible")]
)


# Set up subsets of just the participants enrolled this period
enrolled_master <- to1clean$master[to1clean$master$StudyID %in% enrolled, ]

enrolled_preenroll <- to1clean$preenrollment[to1clean$preenrollment$StudyID 
                                             %in% enrolled, ]



# Set a universal order for the result class variable
result_class_order <- c("Triple Positive", 
                        "TST+, QFT+, Borderline TSPOT",
                        "Dual Pos, TST and QFT",
                        "Dual Pos, TST and TSPOT",
                        "TST+, QFT-, Borderline TSPOT",
                        "Dual Pos, QFT and TSPOT", 
                        "TST-, QFT+, Borderline TSPOT",
                        "Isolated QFT+", 
                        "Isolated TSPOT+", 
                        "Isolated TST+",
                        "TST-, QFT-, Borderline TSPOT",
                        "Triple Negative", 
                        "Inconclusive")




# Cut ages into groups
to1clean$preenrollment$age.grp <- NA
to1clean$preenrollment$age.grp[to1clean$preenrollment$AgeAtEnrollment >= 2 &
                to1clean$preenrollment$AgeAtEnrollment <= 5] <- "2-5"
to1clean$preenrollment$age.grp[to1clean$preenrollment$AgeAtEnrollment >= 6 &
                to1clean$preenrollment$AgeAtEnrollment <= 12] <- "6-12"
to1clean$preenrollment$age.grp[to1clean$preenrollment$AgeAtEnrollment >= 13 &
                to1clean$preenrollment$AgeAtEnrollment <= 25] <- "13-25"
to1clean$preenrollment$age.grp[to1clean$preenrollment$AgeAtEnrollment >= 26 &
                to1clean$preenrollment$AgeAtEnrollment <= 39] <- "26-39"
to1clean$preenrollment$age.grp[to1clean$preenrollment$AgeAtEnrollment >= 40] <- "40+"

# Make it an ordered factor - keeps plots, tables, etc. in sensible order
to1clean$preenrollment$age.grp <- factor(to1clean$preenrollment$age.grp, 
                          levels = c("2-5", "6-12", "13-25", 
                                     "26-39", "40+"),
                          ordered = TRUE
)






```

<h1 class='maintitle'>TBESC TO 1 Results Summary: `r local_facts$site`</h1>
<span class='maintitle'>Report generated at `r format(Sys.time())`</span>


```{r}

# Total enrolled so far (I need a better measure of this)
total_enrolled <- length(enrolled)

```

Total Enrolled to Date: `r total_enrolled`





## Eligibility Criteria
```{r eligcrit}

# Get eligibility criteria from the preenrollments
elig.melt <- melt(enrolled_preenroll, 
                  id.var = "StudyID", 
                  measure.var = c("CloseContact", "ForeignBorn", "LtbiPrevalence",
                                  "HighRisk", "HivPositive")
)

# Convert this weird numeric scheme to logical
elig.melt$valuel <- !is.na(elig.melt$value)

# Cast and summarize
elig.wide <- ddply(elig.melt, .var = "variable", .fun = summarise, n = sum(valuel))


# Add a row for "multiple" criteria
ncrit <- ddply(elig.melt, .var = "StudyID", .fun = summarise, n = sum(valuel))

multicrit <- data.frame(variable = "Multiple",
                        n = nrow(ncrit[ncrit$n > 1, ]))
                            
crit.table <- rbind(elig.wide, multicrit)

crit.table$percent <- round(crit.table$n / nrow(enrolled_preenroll), 4) * 100

levels(crit.table$variable) <- c("Close Contact", "Foreign-born", "Local LTBI Prevalence > 25%",
                                 "30+ Days in a High-Risk Country", "HIV+", "Multiple Criteria")

names(crit.table) <- c("Criterion", "Number Enrolled", "Percent of Enrollees")




```

```{r eligcrit_display, result='asis'}

dfprintif(crit.table)

```




## Gender
```{r gender}

# Summarize
gender <- count(enrolled_preenroll, "Gender")

gender$Gender[gender$Gender %in% 1] <- "Female"
gender$Gender[gender$Gender %in% 2] <- "Male"

# Add percentage
gender$percent <- round(gender$freq / sum(gender$freq), 4) * 100

# Pretty names
names(gender) <- c("Gender", "Number Enrolled", "Percent of Enrollees")


```


```{r gender_display}

dfprintif(gender)

```



## Participant Ages
```{r ages, fig.width=10}

ggplot(enrolled_preenroll, aes(x = AgeAtEnrollment, y = ..count../sum(..count..))) +
    geom_bar(binwidth = 2) +
    scale_y_continuous(labels = percent_format()) +
    labs(x = "Age at Enrollment", y = "Percent of Enrollees") +
    theme_tbesc


```

Median Age of Participants: `r median(enrolled_preenroll$AgeAtEnrollment, na.rm = TRUE)`



## Participants by Country of Origin
```{r,results="hide",fig.width = 8, fig.height = 8}

# Tally up participants by their country of birth
birthco <- arrange(count(enrolled_master, var = "BirthCountry"), 
                   freq, BirthCountry)

# BirthCountry = "" or NA indicates a pre-enrolled person
birthco$BirthCountry[birthco$BirthCountry %in% c("", NA)] <- "Pre-enrolled"

# Make BirthCountry a factor - this ensures descending order of the bars
birthco$BirthCountry <- factor(birthco$BirthCountry,  
                               levels = birthco$BirthCountry)



# Plot it
ggplot(birthco, aes(x = BirthCountry, weight = freq)) +
  geom_histogram() +
  coord_flip() +
  labs(title = "Enrollment by country of birth",
       x = "Country of Birth",
       y = "Number Enrolled") +
  theme_tbesc

```



## Participants' Languages

```{r partlang}

# Convert all langauges to lower case for counting
enrolled_master$interplang <- gsub(
    x = tolower(enrolled_master$InterpreterLanguage),
    pattern = "(^\\w)(.*$)",
    replacement = "\\U\\1\\E\\2",
    perl = TRUE)
            

# A blank means no interpreter was needed (right?)
#count(to1clean$master, vars = c("InterpreterNeeded", "InterpreterLanguage"))

# Recode blanks accordingly
enrolled_master$interplang[enrolled_master$interplang %in% ""] <- "No Interpreter"



# Aggregate
lang <- arrange(count(enrolled_master, var = "interplang"), 
                   freq, interplang)



# Make interplang a factor - this ensures descending order of the bars
lang$interplang <- factor(lang$interplang,  
                          levels = lang$interplang)

ggplot(lang, aes(x = interplang, weight = freq)) +
    geom_histogram() +
    coord_flip() +
    labs(title = "Enrollment by language of interpretation",
         x = "Language",
         y = "Number Enrolled") +
    theme_tbesc



```



## Semiannual Follow-Ups By Language

```{r fu_lang}


# Merge language onto the follow-ups
fu_lang <- merge(x = subset(to1clean$followupfortb,
                            subset = is.na(ParticipantLost),
                            select = c("StudyID", "VisitDate")),
                 y = enrolled_master[c("StudyID", "interplang")],
                 by = "StudyID",
                 all.x = TRUE)

# Aggregate
fu_lang_count <- arrange(count(fu_lang, "interplang"),
                         freq, interplang)

# Make interplang a factor - this ensures descending order of the bars
fu_lang_count$interplang <- factor(fu_lang_count$interplang,  
                                   levels = fu_lang_count$interplang)


ggplot(fu_lang_count, aes(x = interplang, weight = freq)) +
    geom_histogram() +
    coord_flip() +
    labs(title = "Follow-ups by language of interpretation",
         x = "Language",
         y = "Number of Follow-ups") +
    theme_tbesc



```



## Test Results

```{r,results='hide'}
##################################
# Summarize participants' individual test results
testres <- compile_results(to1clean)

testres_enrolled <- testres[testres$StudyID %in% enrolled, ]


# Cast all results into a wide table - keep only the results of
# participants enrolled in this period
testwide <- dcast(
                melt(testres_enrolled,
                     id.var = "StudyID", 
                     measure.var = c("tst", "qft", "tspot")),
                variable ~ value,
                fun.agg = length
)

# Tidy up for printing
# Prettier test names
testwide$variable <- toupper(as.character(testwide$variable))

# Better column order
testwide <- testwide[ , c("variable", "Negative", "Positive", 
                          "Borderline", "Indeterminate")]


# Pretty names
names(testwide)[names(testwide) %in% "variable"] <- "Test"


##################################
# Individual test results - proportions
testwide.prop <- dcast(
                     melt(testres_enrolled,
                          id.var = "StudyID", 
                          measure.var = c("tst", "qft", "tspot")),
                     variable ~ value,
                     fun.agg = function(x) { length(x) / nrow(testres_enrolled) * 100}
)

# Tidy up for printing
# Prettier test names
testwide.prop$variable <- toupper(as.character(testwide.prop$variable))

# Better column order
testwide.prop <- testwide.prop[ , c("variable", "Negative", "Positive", 
                                    "Borderline", "Indeterminate")]


# Pretty names
names(testwide.prop)[names(testwide.prop) %in% "variable"] <- "Test"



##################################
# Aggregate by result class
testagg <- count(testres_enrolled, var = c("result_class"))

# Make a factor of result_class by descending frequency for plotting later on
testres_enrolled$plot_class <- factor(testres_enrolled$result_class,
                                      levels = arrange(testagg, freq)$result_class)



##################################
# Make a factor of result_class for the table - this is more a logical ordering:
# Triple-neg, triple-pos, isolated-pos, isolated-neg
testagg$result_class <- factor(testagg$result_class,
                               levels = c("Triple Negative",
                                          "Triple Positive",
                                          "Isolated TST+",
                                          "Isolated QFT+", 
                                          "Isolated TSPOT+",
                                          "Dual Pos, QFT and TSPOT", 
                                          "Dual Pos, TST and TSPOT", 
                                          "Dual Pos, TST and QFT",
                                          "TST-, QFT-, Borderline TSPOT",
                                          "TST+, QFT-, Borderline TSPOT",
                                          "TST-, QFT+, Borderline TSPOT",
                                          "TST+, QFT+, Borderline TSPOT",
                                          "Inconclusive")
)

```


Count:

```{r}

# Number positive on each test
dfprint(testwide)

```

Proportion:

```{r}
# Proportion positive on each test
dfprint(testwide.prop)

```




```{r, results="hide",fig.width = 8, fig.height = 8}


ggplot(testres_enrolled, aes(x = plot_class)) +
  geom_bar() +
  coord_flip() +
  labs(x = "Number of Participants", y = "Result") +
  theme_tbesc

```


```{r}

# Add a proportion-of-enrolled indicator
testagg$classprop <- round(testagg$freq / nrow(testres_enrolled) * 100, 2)

# Pretty names
names(testagg) <- c("Results", "Freq", "Proportion")


dfprint(arrange(testagg, Results))

```



## Test Positivity by Age Group

```{r posbyagegrp}


# Add age and age group to the test results
testage <- merge(x = testres_enrolled,
                 y = to1clean$preenrollment[ , 
                         c("StudyID", "AgeAtEnrollment", "age.grp")],
                 all.x = TRUE)


# Aggregate by age
testage.agg <- ddply(testage, .var = "age.grp", .fun = summarise,

    n = length(StudyID),
    per.tstpos = sum(tst %in% "Positive") / n,
    per.qftpos = sum(qft %in% "Positive") / n,
    per.tspotpos = sum(tspot %in% "Positive") / n,
    per.tspotbord = sum(tspot %in% "Borderline") / n
                     
)

# Melt it for plotting
testage.melt <- melt(testage.agg, id.vars = c("age.grp", "n"))

# Pretty labels
levels(testage.melt$variable) <- c("TST Positive", 
                                   "QFT Positive",
                                   "TSPOT Positive",
                                   "TSPOT Borderline")


ggplot(testage.melt, 
       aes(x = age.grp, 
           y = value,
           color = variable,
           group = variable)) +
    geom_point(aes(size = n, shape = variable)) +
    geom_line(aes(linetype = variable)) +
    labs(x = "Age at Enrollment", y = "Percent with Result") +
    scale_y_continuous(labels = percent) +
    scale_size_continuous("# of Participants") +
    scale_color_manual("Test", values = brewer.pal(n = 4, name = "Set1")) +
    scale_linetype_discrete("Test") +
    scale_shape_discrete("Test") +
    theme_tbesc

```



## Test Results by Age Group



```{r resultgroupsbyage}
            
# Aggregate results   
ageres <- ddply(testage, .var = "age.grp", .fun = function(x) {
    
    # Aggregate results within the age group
    rescount <- count(x, var = "result_class")
    
    # Calculate each result's percent of the age group
    rescount$per <- round(rescount$freq / nrow(x) * 100, 1)
    
    # Return
    rescount

})


# Make result_class an ordered factor
ageres$result_class <- factor(ageres$result_class,
                              levels = result_class_order)


ggplot(ageres, aes(x = age.grp, weight = per, fill = result_class)) +
    geom_bar(color = "black") +
    labs(x = "Age at Enrollment", y = "Percent of Participants") +
    coord_cartesian(y = c(-1, 60)) +
    scale_y_continuous(breaks = seq(0, 100, 10)) +
    scale_fill_manual("Result Group",
                      values = c(rev(brewer.pal(n = 11, name = "Paired")), 
                                 NA, NA)) +
    theme_tbesc +
    guides(fill = guide_legend(reverse = TRUE))




```



### How much is increasing positivity driven by the triple-positives?

```{r notrippos}

# But actually just reordering the factor and making them invisible
ageres$notrippos <- factor(ageres$result_class,
                           levels = result_class_order[c(2:7, 1, 8, 9)],
                           ordered = TRUE)


ggplot(ageres, aes(x = age.grp, weight = per, fill = notrippos)) +
    geom_bar(color = "black") +
    labs(x = "Age at Enrollment", y = "Percent of Participants") +
    coord_cartesian(y = c(-1, 60)) +
    scale_y_continuous(breaks = seq(0, 100, 10)) +
    scale_fill_manual("Result Group",
                      values = c(rev(brewer.pal(n = 11, 
                                                 name = "Paired")[-7]), 
                                 "honeydew", NA, NA)) +
    theme_tbesc +
    guides(fill = guide_legend(reverse = TRUE))




```


### Selected Result Groups


```{r selected_results_by_age}

# Subset to just the triple-positives, TST-IGRA+, and TST+IGRA-
# and re-aggregate
testage$pos_result_group <- "Other"
testage$pos_result_group[testage$result_class %in% 
                         "Triple Positive"] <- "Triple Positive"
testage$pos_result_group[testage$result_class %in% 
                         "Dual Pos, QFT and TSPOT"] <- "Dual Pos, QFT and TSPOT"
testage$pos_result_group[testage$result_class %in% 
                         "Isolated TST+"] <- "Isolated TST+"
    
testage$pos_result_group <- factor(testage$result_class)
levels(testage$pos_result_group) <- c("Dual Pos, QFT and TSPOT",
                                      NA, NA, NA, NA, NA,
                                      "Isolated TST+",
                                      NA, 
                                      "Triple Positive",
                                      NA, NA, NA, NA)



ageres_pos <- ddply(testage, .var = "age.grp", .fun = function(x) {

    # Aggregate results within the age group
    rescount <- with(x, as.data.frame(table(pos_result_group)))

    # Calculate each result's percent of the age group
    rescount$per <- rescount$Freq / nrow(x)

    # Return
    rescount

})


ggplot(subset(ageres_pos, !is.na(pos_result_group)),
       aes(x = age.grp, y = per, color = pos_result_group)) +
    geom_point(aes(size = Freq, shape = pos_result_group)) +
    geom_line(aes(group = pos_result_group),
              size = 1.2) +
    labs(x = "Age at Enrollment", 
         y = "Percent with Result",
         color = "Result",
         size = "# of Participants",
         shape = "Result",
         linetype = "Result") +
    scale_y_continuous(labels = percent) +
    scale_color_manual(values = brewer.pal(n = 4, name = "Set1")) +
    scale_size_continuous(range = c(4, 10)) +
    theme_tbesc +
    guides(color = guide_legend(override.aes = list(size = 8, linetype = 0)))



```



```{r nar_age_table}

dfprint(count(testage, "age.grp"))


```

<br>

<br>


```{r result_age_table}

# Cast ageres into a more suitable table format
ageres.melt <- melt(ageres,
                    id.var = c("result_class", "age.grp"),
                    measure.var = c("freq", "per"))

ageres.wide <- dcast(ageres.melt,
                     result_class ~ age.grp + variable,
                     value.var = "value")

# Convert N columns to integers
ageres.wide$`2-5_freq` <- as.character(ageres.wide$`2-5_freq`)
ageres.wide$`6-12_freq` <- as.character(ageres.wide$`6-12_freq`)
ageres.wide$`13-25_freq` <- as.character(ageres.wide$`13-25_freq`)
ageres.wide$`26-39_freq` <- as.character(ageres.wide$`26-39_freq`)
ageres.wide$`40+_freq` <- as.character(ageres.wide$`40+_freq`)


# Prettier names
names(ageres.wide) <- c("Results",
                        "Ages 2-5: N", "Ages 2-5: %",
                        "Ages 6-12: N", "Ages 6-12: %",
                        "Ages 13-25: N", "Ages 13-25: %",
                        "Ages 26-39: N", "Ages 26-39: %",
                        "Ages 40+: N", "Ages 40+: %")


dfprintif(ageres.wide)

```




## Test Results by BCG Status

```{r resultsbybcg}

# Add BCG status to the results
testbcg <- merge(x = testage,
                 y = to1clean$medicalhistory[c("StudyID", "BcgVaccine")],
                 by = "StudyID",
                 all.x = TRUE)


# Order BcgVaccine by making it a factor
testbcg$BcgVaccine <-  factor(testbcg$BcgVaccine,
                              levels = c("Yes", "No", "Don't Know/Refused"),
                              ordered = TRUE)


# Reorder result_class, too
testbcg$result_class <- factor(testbcg$result_class,
    levels = c("Triple Negative", 
               "Isolated TST+",
               "Dual Pos, TST and QFT",
               "Dual Pos, TST and TSPOT",
               "Isolated QFT+",
               "Isolated TSPOT+",
               "Dual Pos, QFT and TSPOT",
               "Triple Positive",
               "Inconclusive"),
    ordered = TRUE
)
               


ggplot(testbcg, aes(x = result_class)) +
    geom_bar(aes(fill = BcgVaccine), position = "fill", color = "black") +
    labs(x = "Result Class",
         y = "Proportion of Participants in Result Class") +
    scale_fill_manual("BCG Status", values = c("#1f78b4", "#fb9a99", "#a6cee3")) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 50, hjust = 1))



```


## Quantitative Test Results
```{r quantres}


# Set up a color scale for results
resultcolors <- c("#F03B20", "#1F78B4", "#31A354", "grey", "grey", "grey")

names(resultcolors) <- c("Negative", "Positive", 
                         "Borderline",
                         "Indeterminate", "Invalid", "Test Not Performed")


# Set up result factors for nicer ordering
to1clean$qft$result.ord <- factor(to1clean$qft$result,
                                  levels = c("Negative", "Positive",
                                             "Indeterminate"))


to1clean$tspot$result.ord <- factor(to1clean$tspot$result,
                                   levels = c("Negative", "Borderline",
                                              "Positive", "Invalid",
                                              "Test Not Performed"))


ggplot(to1clean$skintest, aes(x = indur_mm)) +
    geom_histogram(aes(fill = result), binwidth = 1) +
    geom_vline(xintercept = 10, color = "red") +
    labs(x = "TST Induration (mm)", 
         y = "Frequency",
         title = "TST Induration Sizes") +
    scale_x_continuous(breaks = seq(0, 100, 5)) +
    scale_fill_manual("Result", values = resultcolors) +
    theme_bw()


ggplot(to1clean$qft, aes(x = tbnil.num)) +
    geom_histogram(aes(fill = result.ord), binwidth = 0.05) +
    geom_vline(xintercept = 0.35, color = "red") +
    labs(x = "TB - Nil (IU/mL)", 
         y = "Frequency",
         title = "Quantiferon TB - Nil Results") +
    scale_x_continuous(breaks = c(seq(-10, 0, 1), 0.35, seq(1, 10, 1))) +
    scale_fill_manual("Result", values = resultcolors) +
    theme_bw()


ggplot(to1clean$tspot, aes(x = tbnil.max)) +
    geom_histogram(aes(fill = result.ord), binwidth = 1) +
    geom_vline(xintercept = c(5, 8), color = "red") +
    labs(x = "Maximum TB - Nil Result (spots)", 
         y = "Frequency",
         title = "TSPOT TB - Nil Results") +
    scale_x_continuous(breaks = c(seq(-20, 5, 5), 8, seq(10, 100, 5))) +
    scale_fill_manual("Result", values = resultcolors) +
    theme_bw()


```



## TST Induration Size by BCG Status

```{r tst_by_bcg, res = '300px'}

# Add BCG status to TST results
tst_bcg <- merge(x = to1clean$skintest,
                 y = to1clean$medicalhistory[c("StudyID", "BcgVaccine")],
                 all.x = TRUE,
                 by = "StudyID")



ggplot(tst_bcg, aes(x = BcgVaccine, y = indur_mm)) +
    geom_violin(fill = alpha(colorwheel[3], 0.1), scale = "area") +
    geom_boxplot(width = 0.1, fill = alpha(colorwheel[3], 0.5)) +
    stat_summary(fun.y = "median", geom = "point") +
    labs(x = "BCG Status",
         y = "TST Induration (mm)") +
    theme_bw()


```



## Quantitative Test Results By Age
```{r quantres_age}



# Set up defined shapes for the results
resultshapes <- c(15, 3, 18, 13, 13, 13)

names(resultshapes) <- c("Negative", "Positive", 
                         "Borderline",
                         "Indeterminate", "Invalid", "Test Not Performed")


# Merge ages onto quantitative results
tstage <- merge(x = to1clean$skintest,
                y = to1clean$preenrollment[c("StudyID", 
                                             "AgeAtEnrollment", 
                                             "age.grp")],
                by = "StudyID",
                all.x = TRUE)

qftage <- merge(x = to1clean$qft,
                y = to1clean$preenrollment[c("StudyID",
                                             "AgeAtEnrollment",
                                             "age.grp")],
                by = "StudyID",
                all.x = TRUE)

tspotage <- merge(x = to1clean$tspot,
                  y = to1clean$preenrollment[c("StudyID",
                                               "AgeAtEnrollment",
                                               "age.grp")],
                  by = "StudyID",
                  all.x = TRUE)




ggplot(tstage, aes(x = age.grp, y = indur_mm)) +
    geom_boxplot(fill = colorwheel[2]) +
    geom_hline(yintercept = 10, color = "red") +
    labs(x = "Age At Enrollment", 
         y = "Induration (mm)",
         title = "TST Induration Sizes by Age Group") +
    scale_color_manual("Result", values = resultcolors) +
    scale_shape_manual("Result", values = resultshapes) +
    theme_bw()


ggplot(qftage, aes(x = age.grp, y = tbnil.num)) +
    geom_boxplot(fill = colorwheel[2]) +
    geom_hline(yintercept = 0.35, color = "red") +
    labs(x = "Age At Enrollment", 
         y = "TB - Nil Results (IU/mL)",
         title = "Quantiferon TB - Nil Results by Age Group") +
    scale_y_continuous(breaks = c(seq(-10, 0, 1), 0.35, seq(1, 10, 1))) +
    scale_color_manual("Result", values = resultcolors) +
    scale_shape_manual("Result", values = resultshapes) +
    theme_bw()



ggplot(tspotage, aes(x = age.grp, y = tbnil.max)) +
    geom_boxplot(fill = colorwheel[2]) +
    geom_hline(yintercept = c(5, 8), color = "red") +
    labs(x = "Age At Enrollment", 
         y = "Maximum TB - Nil Result (spots)",
         title = "TSPOT TB - Nil Results by Age Group") +
    scale_y_continuous(breaks = c(seq(-20, 5, 5), 8, seq(10, 100, 5))) +
    scale_color_manual("Result", values = resultcolors) +
    scale_shape_manual("Result", values = resultshapes) +
    theme_bw()



```



```{r mitonilbyage}

qftage.m <- melt(qftage,
                 id.vars = c("StudyID", "age.grp"),
                 measure.vars = c("nil.num", "mito.num"))

tspotage.m <- melt(tspotage,
                   id.vars = c("StudyID", "age.grp"),
                   measure.vars = c("nil.num", "mito.num"))


# Nicer labels
levels(qftage.m$variable) <- c("Nil", "Mitogen")
levels(tspotage.m$variable) <- c("Nil", "Mitogen")



ggplot(qftage.m, aes(x = age.grp, y = value)) +
    geom_boxplot(aes(fill = variable)) +
    geom_hline(yintercept = 0.35, color = "red") +
    labs(x = "Age At Enrollment", 
         y = "Quantitative Result (IU/mL)",
         title = "Quantiferon Nil and Mitogen Results by Age") +
    scale_y_continuous(breaks = c(seq(-10, 0, 1), 0.35, seq(1, 10, 1))) +
    scale_fill_manual("Result", values = unname(resultcolors)) +
    theme_bw()



ggplot(tspotage.m, aes(x = age.grp, y = value)) +
    geom_jitter(aes(color = variable)) +
    geom_hline(yintercept = c(5, 8), color = "red") +
    labs(x = "Age At Enrollment", 
         y = "Quantitative Result (spots)",
         title = "TSPOT Nil and Mitogen Results by Age") +
    scale_y_continuous(breaks = c(seq(-20, 5, 5), 8, seq(10, 100, 5))) +
    scale_color_manual("Result", values = unname(resultcolors)) +
    theme_bw()



```



```{r qftmodel, results='hide'}


qftage.meltall <- melt(qftage,
                       id.vars = c("StudyID", "AgeAtEnrollment"),
                       measure.vars = c("nil.num", "tb.num", "mito.num"))


# ggplot(qftage.meltall, aes(x = value, y = ..count..)) +
#     geom_histogram(binwidth = 0.5)


# With TB
qftmod.var <- lm(value ~ variable, data = qftage.meltall)
summary(qftmod.var)
# qplot(resid(qftmod.var), binwidth = 0.50)

qftmod.age <- lm(value ~ variable * AgeAtEnrollment, data = qftage.meltall)
summary(qftmod.age)
# qplot(resid(qftmod.age), binwidth = 0.50)

AIC(qftmod.var)
AIC(qftmod.age)



# Without TB
qftage.notb <- subset(qftage.meltall, variable %in% c("nil.num", "mito.num"))

qftnotb.var <- lm(value ~ variable, data = qftage.notb)
summary(qftnotb.var)
# qplot(resid(qftnotb.var), binwidth = 0.50)

qftnotb.age <- lm(value ~ variable * AgeAtEnrollment, data = qftage.notb)
summary(qftnotb.age)
# qplot(resid(qftnotb.age), binwidth = 0.50)

AIC(qftnotb.var)
AIC(qftnotb.age)

```


```{r tspotmodel, results='hide'}


tspotage.meltall <- melt(tspotage,
                       id.vars = c("StudyID", "AgeAtEnrollment"),
                       measure.vars = c("nil.num", "mito.num",
                                        "panel_a.num", "panel_b.num"))


# ggplot(tspotage.meltall, aes(x = value, y = ..count..)) +
#     geom_histogram(binwidth = 0.5)


# Truncate values to 20
#tspotage.meltall$value[tspotage.meltall$value > 20] <- 20

# With TB
tspotmod.var <- lm(value ~ variable, data = tspotage.meltall)
summary(tspotmod.var)
# qplot(resid(tspotmod.var), binwidth = 0.50)

tspotmod.age <- lm(value ~ variable * AgeAtEnrollment, data = tspotage.meltall)
summary(tspotmod.age)
# qplot(resid(tspotmod.age), binwidth = 0.50)

AIC(tspotmod.var)
AIC(tspotmod.age)



# Without TB
tspotage.notb <- subset(tspotage.meltall, variable %in% c("nil.num", "mito.num"))

tspotnotb.var <- lm(value ~ variable, data = tspotage.notb)
summary(tspotnotb.var)
# qplot(resid(tspotnotb.var), binwidth = 0.50)

tspotnotb.age <- lm(value ~ variable * AgeAtEnrollment, data = tspotage.notb)
summary(tspotnotb.age)
# qplot(resid(tspotnotb.age), binwidth = 0.50)

AIC(tspotnotb.var)
AIC(tspotnotb.age)

```


## Test Results Over the Enrollment Period

```{r tests_overtime}

# Aggregate by enrollment month
testres_enrolled$EnrollMonth <- format(testres_enrolled$EnrollDate, "%Y-%m")


monthmelt <- melt(subset(testres_enrolled, EnrollDate >= as.Date("2012-11-01")),
                  id.vars = "EnrollMonth",
                  measure.vars = c("tst", "qft", "tspot"))


# Aggregate
monthcount <- ddply(monthmelt,
                    .var = c("EnrollMonth", "variable"),
                    .fun = summarise,
                    
    n = length(value),
    per.pos = sum(value %in% "Positive") / n,
    per.bord = sum(value %in% "Borderline") / n,
    per.ind = sum(value %in% "Indeterminate") / n
                    
)
    

# Tidy up for plot
levels(monthcount$variable) <- c("TST", "QFT", "TSPOT")


ggplot(monthcount, aes(x = EnrollMonth, 
                      y = per.pos,
                      color = variable)) +
    geom_point(aes(size = n)) +
    geom_line(aes(group = variable, linetype = variable)) +
    expand_limits(y = 0) +
    scale_y_continuous(labels = percent) +
    scale_color_manual("Test", 
                       values = colorwheel[1:3]) +
    scale_linetype_discrete("Test") +
    scale_size_continuous("# Enrolled") +
    labs(x = "Enrollment Month", y = "Percent Positive") +
    theme_tbesc +
    theme(axis.text.x = element_text(angle = 70, hjust = 1))
    
    




```

## Treatment Overview

```{r tx_regimens}

# Get the tx info from DMS - probably a way to refactor with the (older) code
# in Treatment Outcomes By Test Result below
# regimens <- subset(to1clean$ltbi,
#                    subset = OfferTreatment,
#                    select = c("StudyID", "OfferDate",
#                               "OfferTreatment", "AcceptTreatment",
#                               "plan", "TreatmentComplete")
# )

regimens <- subset(summarize_tx(to1clean), subset = OfferTreatment)


n_offered <- sum(regimens$OfferTreatment, na.rm = TRUE)
n_accepted <- sum(regimens$AcceptTreatment, na.rm = TRUE)
per_accepted <- calc_percent(n_accepted, n_offered, digits = 0)


```


A total of `r n_offered` participants have been
offered treatment; `r n_accepted` (`r per_accepted`%) accepted treatment.


```{r regimen_table}




# Aggregate the regimens
regimen_table <- ddply(subset(regimens, AcceptTreatment), 
                       .var = "plan", 
                       .fun = summarise,
                       
    # Number and percent of participants started on this regimen
    n_reg = length(StudyID),
    per_reg = calc_percent(n_reg, n_accepted, digits = 1),
                       
    # Number of participants eligible to complete - count all who have had
    # enough_time, as well as any who have already completed
    n_eligible = sum(TreatmentComplete | enough_time, na.rm = TRUE),
                       
    # Number and percent of the eligible who have completed
    n_completed = sum(TreatmentComplete, na.rm = TRUE),
    per_completed = calc_percent(n_completed,
                                 n_eligible,
                                 digits = 1)
                       
)

# Nicer names
names(regimen_table) <- c("Regimen", 
                          "N Started", "% of Starts", 
                          "N Eligible to Complete",
                          "N Completed", "% Completed")


dfprintif(arrange(regimen_table, desc(`N Started`)))


```

Note: Completion rates are calculated using participants who have had sufficient
time to complete following CDC guidelines (52 weeks for Daily INH,
26 weeks for Daily RIF, and
16 weeks for Weekly INH/Rifapentine).

Participants who have completed treatment are 
included regardless of time on regimen.




## Regimen Starts Over Time

```{r reg_over_time}

# Aggregate by... quarter?
regimens$offer_qtr <- paste0(
    format(regimens$OfferDate, "%Y"),
    " Q",
    (as.numeric(format(regimens$OfferDate, "%m")) + 2) %/% 3)


# Convert plan to a factor so that plans get at least a zero-count in all quarters
regimens$plan.f <- factor(regimens$plan)

reg_over_time <- ddply(regimens, .var = "offer_qtr", .fun = function(x) {
    
    # Aggregate by plan
    plan_count <- with(x, as.data.frame(table(plan.f)))
    
    plan_count$prop <- plan_count$Freq / sum(plan_count$Freq)
    
    plan_count
    
})


    
ggplot(reg_over_time, aes(x = offer_qtr, y = prop, color = plan.f)) +
    geom_point(aes(size = Freq)) +
    geom_line(aes(group = plan.f), size = 1.2) +
    labs(x = "Quarter Offered",
         y = "Percent of Participants Who Initiated Treatment") +
    scale_color_brewer("Tx Regimen", palette = "Paired") +
    scale_y_continuous(labels = percent) +
    scale_size_continuous("Number Starting Tx", range = c(3, 10)) +
    theme_bw() + theme(legend.key.width = unit(1, "in"))




```


## Treatment Outcomes By Test Result


```{r tx_outcomes}


# Merge treatment info onto test results
tx_test <- merge(x = testres_enrolled,
                 y = regimens,
                 by = "StudyID",
                 all.x = TRUE)


# Summarize by single test results
tx_test_melt <- melt(tx_test,
                     id.var = c("StudyID", "OfferTreatment",
                                "AcceptTreatment", "enough_time",
                                "TreatmentComplete"),
                     measure.var = c("tst", "qft", "tspot")
)

result_tx  <- ddply(tx_test_melt, 
                    .var = c("variable", "value"), 
                    .fun = summarise,
                    
    # Number of participants positive by this test
    n = length(StudyID),
                    
    # Number and % of test-pos participants offered treatment
    n_offered = sum(OfferTreatment, na.rm = TRUE),
    per_offered = round(n_offered / n * 100, 2),
                    
    # Number and % of test-pos participants who were
    # offered treatment and accepted
    n_accepted = sum(AcceptTreatment, na.rm = TRUE),
    per_accepted = round(n_accepted / n_offered * 100, 2),
                    
    # Number of participants eligible to complete - count all who have had
    # enough_time, as well as any who have already completed
    n_eligible = sum(TreatmentComplete | enough_time, na.rm = TRUE),

    # Number and percent of the eligible who have completed
    n_completed = sum(TreatmentComplete, na.rm = TRUE),
    per_completed = calc_percent(n_completed,
                                 n_eligible,
                                 digits = 1)
                    
)


# Subset to just positive tests
pos_tx <- subset(result_tx, value %in% "Positive")

# Relabel variable
pos_tx$variable <- c("TST+", "QFT+", "TSPOT+")




# Summarize by result class                 
result_class_tx  <- ddply(tx_test, 
                    .var = "result_class", 
                    .fun = summarise,
                    
    # Number of participants with this result class
    n = length(StudyID),

    # Number and % of this class who were offered treatment
    n_offered = sum(OfferTreatment, na.rm = TRUE),
    per_offered = round(n_offered / n * 100, 2),

    # Number and % of this class who were offered treatment and accepted
    n_accepted = sum(AcceptTreatment, na.rm = TRUE),
    per_accepted = round(n_accepted / n_offered * 100, 2),

    # Number of participants eligible to complete - count all who have had
    # enough_time, as well as any who have already completed
    n_eligible = sum(TreatmentComplete | enough_time, na.rm = TRUE),

    # Number and percent of the eligible who have completed
    n_completed = sum(TreatmentComplete, na.rm = TRUE),
    per_completed = calc_percent(n_completed,
                                 n_eligible,
                                 digits = 1)

)


# Make result class a factor
result_class_tx$result_class <- factor(result_class_tx$result_class,
    levels = result_class_order,
    ordered = TRUE)

# Replace NaNs with -- for tidier display
result_class_tx$per_accepted[is.nan(result_class_tx$per_accepted)] <- "--"
result_class_tx$per_completed[is.nan(result_class_tx$per_completed)] <- 
    "--"


# Replace n_accept with zero when no one was offered,
# and n_complete with zero when no one accepted
result_class_tx$n_accepted[result_class_tx$n_offered %in% 0] <- "--"
result_class_tx$n_completed[result_class_tx$n_accepted %in% 0] <- "--"



```

```{r print_pos_tx}


# Prettier names
names(pos_tx) <- c("Result", "drop", 
                   "N", 
                   "N Offered", "% Offered",
                   "N Accepted", "% Accepted",
                   "N Eligible to Complete",
                   "N Completed", "% Completed")


dfprintif(pos_tx[ , names(pos_tx)[!names(pos_tx) %in% "drop"]])



```

Note: Completion rates are calculated using participants who have had sufficient
time to complete following CDC guidelines (52 weeks for Daily INH,
26 weeks for Daily RIF, and
16 weeks for Weekly INH/Rifapentine).

Participants who have completed treatment are 
included regardless of time on regimen.




## Treatment Outcomes By Result Group


```{r print_class_tx}


# Prettier names
names(result_class_tx) <- c("Result Class", "N", 
                            "N Offered", "% Offered",
                            "N Accepted", "% Accepted",
                            "N Eligible to Complete",
                            "N Completed", "% Completed")


dfprintif(arrange(result_class_tx, `Result Class`))



```

Note: Completion rates are calculated using participants who have had sufficient
time to complete following CDC guidelines (52 weeks for Daily INH,
26 weeks for Daily RIF, and
16 weeks for Weekly INH/Rifapentine).

Participants who have completed treatment are 
included regardless of time on regimen.




### Semiannual Follow-up Results

```{r semiannual_fu}

# Calculate the number of participants eligible for follow-ups
fu_eligible <- subset(to1clean$master, 
                      subset = ParticipantStatus %in% c("LTBI Test +",
                                                        "LTBI Treatment",
                                                        "LTBI Not Treated"),
                      select = c("StudyID", "EnrollDate", 
                                 "BirthCountry", "InterpreterLanguage")
)

# How have at least one semiannual follow-up record?
fu_eligible$recorded <- fu_eligible$StudyID %in% to1clean$followupfortb$StudyID

# How many are lost to follow-up?
fu_eligible$lost <- fu_eligible$StudyID %in%
    to1clean$followupfortb$StudyID[!is.na(to1clean$followupfortb$ParticipantLost)]


# Aggregate
fu_stats <- summarise(fu_eligible,
                      n_eligible = length(StudyID),
                      n_recorded = sum(recorded),
                      n_lost = sum(lost),
                      per_lost = calc_percent(n_lost, n_recorded, digits = 1)
)


# Rename for printing
names(fu_stats) <- c("N Eligible for FU",
                     "N with 1+ Recorded FU",
                     "N Lost",
                     "% Lost")

dfprintif(fu_stats)


```



### Test Results by Country of Origin




### Test Results by Risk Factors








### TB Disease




